Sensor Clusterization in D-optimal Design in Infinite Dimensional Bayesian Linear Inverse Problems

### 2024-09-08 revision/rev3.tex
Typos
Done

\section{Reviewer 3}\label{ref3}
\RC The paper studies Bayesian D-optimal design (that is, the task of
allocating measurements in order to maximise the Kullback-Leibler
divergence between the prior and posterior distribution) in an inverse
regression type model in Hilbert spaces with Gaussian measurements
errors. For Gaussian priors, the paper provides a characterisation of
the optimality criterion and the optimal design. Some conclusions
about the phenomenon of clusterisation of measurements known in
literature are then drawn.

@@

\AR I made sure the repository is functioning and well-documented. See
\url{https://github.com/yairdaon/OED}.

@@

\begin{quote} %% Simulate
  In the proof of Theorem \ref{thm:char} we utilize Lemma
  \ref{lemma:free} to construct D-optimal designs. We implement this
  construction with the goal of testing numerically how prevalent are
  clustered designs. To this end, we would like to generate random
  prior eigenvalues $\lambda_j$, fix $m$ and $k$, find a D-optimal
  $\opt^*\opt$ and then utilize the construction of
  Theorem~\ref{thm:char} and Lemma~\ref{lemma:free} to find $\opt$.
  \newline
  To simplify things, we directly sample rank $\opt^*\opt$. We iterate
  over the number of measurements $m \in \{4,\dots, 24\}$, and for
  every $m$ we then iterate over $k:=\rank \obs^*\obs \in \{2,\dots,
  m-1\}$. For each pair $m,k$ we repeat the following steps $N=5000$
  times:
  \begin{enumerate}
  \item Generate random diagonal $D\in \mathbb{R}^{k\times k}$ with
    entries $\log (d_i) \sim \mathcal{N}(50,15)$ and normalize so that
    $\ttr D = m$.
  \item Conjugate $D$ by a random orthogonal matrix to form a positive
    semi-definite $M := UDU^t \in \mathbb{R}^{k\times k}$. This $M$
    represents $\opt^*\opt$.
  \item Apply the construction of Lemma \ref{lemma:free} to calculate
    $A$ such that $AA^t = M$, where $A$ has unit norm
    columns. I.e.~find the optimal design $\opt$.
  \item Since $A$ corresponds to $\opt$, its columns correspond to
    measurement vectors. We call $A$ "clustered" if $A$ has two or
    more identical columns (up to some numerical precision threshold).
  \end{enumerate}
  We then calculate the fraction of clustered designs of the
  simulations we ran, for each pair $m,k$. Clusterization occurred at
  high rates ($>99.9\%$) whenever $m-k > 1$; see
  Fig.~\ref{fig:sim_AAt}. Hence, in these simulations, clusterization
  is a generic property. However, when $m-k = 1$, clusterization does
  not occur. We do not why this is so and further investigation into
  this phenomenon is out of scope for the current study.
  \newline
  Full results are located in the \texttt{simulations.csv} file within
  the accompanying \href{https://github.com/yairdaon/OED}{repository}.
  Code implementing the experiments described above is located in
  module \texttt{zeros.py} of said repository. Runtime should be less
  than 30 minutes on any reasonably modern laptop (it took 12 minutes
  on the author's laptop).
\end{quote}

\begin{figure}
    \centering
    \includegraphics[height=0.5\textwidth]{../latex/figs/simulations.png}
    \caption{Fraction of clustered $A$ for $AA^t = M$ and $M$
      generated randomly (see text and repository for details on
      generating $M$). It is evident that when $m-k \geq 2$ clusterization
      is ubiquitous, whereas for lower $m-k$ clusterization does not
      occur.}
  \label{fig:sim_AAt}
\end{figure}

@@

\AR I added a discussion on this subject. In the manuscript this
discussion is broken in two since the second part relies on Theorem
\ref{thm:char}. Thus, the first part is in the Introduction and the
second part follows the statement and proof of the abovementioned
theorem. I believe the level of discussion is adequate and view a
deeper analysis of convergence rates as out of scope for the current
study.
\begin{quote}%%Convergence
  Given that D-optimal designs suffer from clusterization, it is
  natural to ask whether D-optimal designs should be pursued at
  all. Other authors have established convergence results for decaying
  measurement error \cite{knapik2011}, space-filling measurements
  \cite{teckentrup2020} and randomly sampled designs
  \cite{nickl2023}. Unfortunately, D-optimal designs --- clustered
  designs in particular --- do not fall into any of these categories.
  \newline
  Luckily, we can expect convergence from D-optimal designs (including
  clustered designs). Convergence holds in the following sense: the
  posterior uncertainty ellipsoid in $\hilo$ will contract to zero
  along every eigenfunction of $\fwd \prcov \fwd^*$. See Section
  \ref{subsub:convergence} for a proof based on Theorem \ref{thm:char}
  --- the main theorem of this manuscript.
  \newline
  Below we fulfill our promise from Section \ref{subsub:implications}
  and prove that the posterior uncertainty ellipsoid in $\hilo$ will
  contract to zero along every eigenfunction of $\fwd \prcov
  \fwd^*$. In our proof we ignore potential problems with conducting
  inference on function spaces which are not unique to D-optimal
  designs (see e.g.~\ref{owhadi2015} for more details).
  \newline
  First, let $\opt_m$ a D-optimal design utilizing $m$
  measurements. Then, an immediate consequence of Theorem
  \ref{thm:char} is that $\rank \opt_m^*\opt_m = k \to \infty$ as the
  number of measurements $m \to \infty$. Moreover, part (5) of Theorem
  \ref{thm:char} ensures that the eigenvalues of the posterior
  covariance equal
  \begin{equation*}
    \theta^{(k)}_i = \left ( \frac{\sum_{j=1}^k \lambda_j^{-1} +
      \sigma^{-2}m}{k} \right )^{-1}, \text{ for $i\leq k$},
  \end{equation*}
  where we explicitly express the dependence of $k$ in the
  superscript. Now note that $\theta^{(k)}_i \leq \lambda_i$ for
  $i\leq k$. Specifically, for fixed $k$, we get $\theta^{(k)}_i \leq
  \lambda_k$ for all $i\leq k$. Since $\lim_{k\to \infty} \lambda_k=
  0$, we conclude that for all $i$, $\lim_{k\to\infty} \theta^{(k)}_i
  = 0$. Since $m\to \infty$ implies $k\to \infty$ we have
  $\lim_{m\to\infty} \theta^{(k)}_i = 0$ for all $i$. Therefore,
  posterior uncertainty decays to zero for all eigenvectors.
\end{quote}

@@

\AR These are two great points. Indeed, the growth of eigenvalues of
the Laplacian is different in higher dimensions. This is why in higher
dimensions we would have to take as prior \(u_0 \sim
\mathcal{N}(\param_0, (-\Delta)^{-\gamma})\), for some \(\gamma >
d/2\) in order to maintain regularity of prior realizations. I decided
not to include a discussion on the choice of priors in higher spatial
dimensions because (1) it would require too much exposition and make
me digress from the main points I wanted to convey and (2) due to
technical difficulties in installing packages for inverse problems
\cite{attia2023pyoed, VillaPetraGhattas16} I had to implement the
inverse problem of the 1D heat equation from scratch, so I decided to
leave out experiments in higher dimensions.

@@

\AR I removed the statement of Theorem 1 from the introduction. It is
Theorem \ref{thm:char} now.

@@

\begin{quote} %% Replication
Clusterization should not be confused with replication. Replication
requires that the experimentalist executes multiple trials under
circumstances that are \emph{nominally identical} \cite[Section
  1.2.4]{morris2011}. Replication is commonly viewed as a beneficial
and even necessary aspect of optimal experimental design
\cite{fisher1949design, morris2011, schafer2001replication}. Similarly
to a design implementing replication, a clustered design reduces the
signal-to-noise ratio of the repeated measurements
\cite{telford2007brief}. The difference is that a clustered design
takes repeated measurements at the expense of other quantities not
measured at all.
\end{quote}

@@

\AR the Mat\'ern-like prior actually arises by using a Laplacian-like
operator as a prior \cite{rue2011}. So a study of this example (with
and without correlated errors) is already included in the paper. I am
afraid implementing another inverse problem is out of scope for the
present paper since I was not able to install common packages for
inverse problems \cite{attia2023pyoed, VillaPetraGhattas16} on my
machine.

@@

\AR Indeed, point evaluations $\delta_x$ are not in any function space
I consider in the paper since my analysis is limited to Hilbert
spaces. Point evaluations can be approximated well in
e.g.~$L^2(\Omega)$ so in my opinion this is not a deal breaker. For
example, we can replace point evaluations by $R(x;t) =
t\mathbb{1}_{-1/2t, 1/2t}(x)$, where $\mathbb{1}$ is an indicator
function. We will then need to change the norm constraints on
measurement vectors from unit to $\|R(\cdot;t\|^2 = \int R^2(x;t)dx =
t$ norm but that does not change the analysis as long as $t$ is
fixed. I barely mentioned this issue in the paper since I think it is
confusing and adds little context. See a footnote I added:

@@

\AR I explain this in more detail in the revised manuscript:

\begin{quote}
  Building on Theorem \ref{thm:char}, we can now give a compelling
  explanation to the measurement clusterization we observed for the
  inverse problem of the heat equation, see Section
  \ref{subsub:clusterization1} below. We also suggest a generic
  explanation for clusterization, see Section
  \ref{subsub:clusterization2}.
  \newline
  Consider $\fwd$ and $\prcov$ from \emph{the inverse problem of the
  heat equation}. As before, we denote the eigenvalues of
  $\fwd\prcov\fwd^*$ by $\lambda_j$. We input these eigenvalues into
  our \emph{generic} model, and find a D-optimal design $\opt$ for our
  generic model using Theorem \ref{thm:char}. In our generic model,
  the measurements we take are best utilized in reducing uncertainty
  for the first $k$ eigenvectors. So, a D-optimal design arising from
  our \emph{generic model} completely avoids measuring eigenvectors
  $k+1$ and above.
  \newline
  Of course, in a real life problem --- such as the inverse problem of
  the 1D heat equation --- it is likely impossible to find measurement
  locations for which all eigenvectors $k+1$ and above are
  zero. However, if the eigenvalues of $\fwd\prcov\fwd^*$ decay
  quickly (recall the square-exponential decay for eigenvalues of the
  1D heat equation in eq.\eqref{eq:decay}), a D-optimal design will
  try to balance measuring a small number (i.e.~$k$) of the leading
  eigenvectors.
  \newline
  The abovementioned balance is explored in
  Fig.~\ref{fig:eigenvectors}. We allow $m=4$ measurements in $\Omega
  = [0,1]$ and observe that D-optimal measurement locations are
  clustered at $x_1 = 0.31$ and $x_2 = 0.69$. Upon close inspection of
  the scaled eigenvectors of $\fwd \prcov \fwd^*$, we first observe
  that eigenvectors $3$ and above have negligible prior
  amplitude. Since we only have $m=4$ measurements at our disposal, we
  interpret these results, following Theorem \ref{thm:char}, as
  implying we should only care about measuring the first and second
  eigenvectors. Then, we note the D-optimal $x_1,x_2$ present a
  compromise between the amplitude of the first and second
  eigenvectors. For example, a measurement at $x=0.5$ would have
  ignored the second eigenvector altogether, since the second
  eigenvector is zero at $x=0.5$.
  \newline
  Now we can understand measurement clusterization for the inverse
  problem of the heat equation. A D-optimal design attempts to measure
  the first $k$ eigenvectors of $\fwd \prcov \fwd^*$. But there may be
  (spatial) limitations on where these $k$ eigenvectors have large
  amplitude. For the inverse problem of the heat equation there are
  two spatial locations that present a good compromise between the
  amplitudes of the first and second eigenvectors, namely $x_1$ and
  $x_2$ --- see Fig.~\ref{fig:eigenvectors}. We have $m=4$
  measurements at our disposal but only two spatial locations that are
  a good compromise between the first and second scaled
  eigenvectors. Thus, clusterization arises as a consequence of the
  pigeonhole principle.
\end{quote}

\begin{figure}\label{fig:eigenvectors}
  \centering
  \includegraphics[width=\textwidth]{../latex/figs/eigenvectors_dst_scaled.png}
  \caption{D-optimal measurement locations ($m=4$ measurements) and
    weighted eigenvectors for finding the initial condition of the 1D
    heat equation. Measurement locations and weighted eigenvectors are
    plotted over the computational domain $\Omega = [0, 1]$
    (x-axis). Measurement clusterization occurs approximately at
    $0.31$ and $0.69$. These two locations are a compromise between
    the magnitudes of the first and second eigenvectors, which are the
    eigenvectors that a D-optimal design aims to measure. Allocating
    $m=4$ measurements into two locations results in clusterization,
    according to the pigeonhole principle.}
  \label{fig:why}
\end{figure}

>>>>

\section{Referee 3}\label{ref3}
\RC The paper studies Bayesian D-optimal design (that is, the task of
allocating measurements in order to maximise the Kullback-Leibler
divergence between the prior and posterior distribution) in an inverse
regression type model in Hilbert spaces with Gaussian measurements
errors. For Gaussian priors, the paper provides a characterisation of
the optimality criterion and the optimal design. Some conclusions
about the phenomenon of clusterisation of measurements known in
literature are then drawn.

@@

\AR I made sure the repository is functioning and well-documented. See
\url{https://github.com/yairdaon/OED}. Furthermore, I added references
to specific modules in the text, see excerpts below.

\begin{quote}
Full results are located in the \texttt{simulations.csv} file within
the accompanying \href{https://github.com/yairdaon/OED}{repository}.
Code implementing the experiments described above is located in module
\texttt{zeros.py} of said repository.
\end{quote}
\begin{quote}
Code generating these movies is located in module \texttt{movies.py}
in the accompanying
\href{https://github.com/yairdaon/OED}{repository}.
\end{quote}
\begin{quote}
The implementation of the D-optimality criterion for this inverse
problem with homogeneous Neumann and Dirichlet boundary conditions can
be found in directory \texttt{src/} and script
\texttt{clusterization.py} in the accompanying
\href{https://github.com/yairdaon/OED}{repository}.
\end{quote}

@@

\begin{quote} %% Simulate
  In the proof of Theorem \ref{thm:char} we utilize Lemma
\ref{lemma:free} to construct D-optimal designs. We implement this
construction with the goal of testing numerically how prevalent are
clustered designs. To this end, we would like to generate random prior
eigenvalues $\lambda_j$, fix $m$ and $k$, find what a D-optimal
$\opt^*\opt$ should be, and then utilize the construction of
Theorem~\ref{thm:char} and Lemma~\ref{lemma:free} to find $\opt$.

To simplify things, we directly generate $\opt^*\opt$. We iterate over
the number of measurements $m \in \{4,\dots, 24\}$, and for every $m$
we then iterate over $k:=\rank \opt^*\opt \in \{2,\dots, m-1\}$. For
each pair $m,k$ we repeat the following steps $N=5000$ times:
\begin{enumerate}
\item Generate random diagonal $D\in \mathbb{R}^{k\times k}$ with
  entries $\log (d_i) \sim \mathcal{N}(50,15)$.
\item Normalize $D$ that $\ttr D = m$.
\item Conjugate $D$ by a random orthogonal matrix to form a positive
  semi-definite $M := UDU^t \in \mathbb{R}^{k\times k}$. This $M$
  represents $\opt^*\opt$.
\item Apply the construction of Lemma \ref{lemma:free} to calculate
  $A$ such that $AA^t = M$, where $A$ has unit norm columns. $A^t$ is
  our optimal design $\opt$.
\item Since $A^t$ corresponds to $\opt$, its columns correspond to
  measurement vectors. We call $A$ "clustered" if $A$ has two or more
  identical columns (up to some numerical precision threshold,
  i.e.~$10^{-5}$).
\end{enumerate}
We then calculate the fraction of clustered designs of the simulations
we ran, for each pair $m,k$. Clusterization occurred at high rates
($>99.9\%$) whenever $m-k > 1$; see Fig.~\ref{fig:sim_AAt}. Hence, in
these simulations, clusterization is a generic property. However, when
$m-k = 1$, clusterization does not occur. We do not why this is so.
\newline
\newline
Full results are located in the \texttt{simulations.csv} file within
the accompanying \href{https://github.com/yairdaon/OED}{repository}.
Code implementing the experiments described above is located in module
\texttt{zeros.py} of said repository. Runtime should be less than 30
minutes on any reasonably modern laptop (it took 12 minutes on the
author's laptop).
\end{quote}

\begin{figure}
    \centering
    \includegraphics[height=0.5\textwidth]{../latex/figs/simulations.png}
    \caption{Fraction of clustered $A$ for $AA^t = M$ and $M$
      generated randomly (see text and repository for details on
      generating $M$). It is evident that when $m-k > 1$ clusterization
      is prevalent, whereas for lower $m-k$ clusterization is not.}
  \label{fig:sim_AAt}
\end{figure}

@@

\AR I added a discussion on this subject. In the manuscript this
discussion is broken in two since the second part relies on Theorem
\ref{thm:char}. Thus, the first part is in the Introduction and the
second part follows the statement and proof of the abovementioned
theorem.
\begin{quote}%%Convergence
  Clusterization in D-optimal designs raises the question of whether
  D-optimal designs should be pursued at all. While other authors have
  established convergence results for decaying measurement error
  \cite{knapik2011}, space-filling measurements \cite{teckentrup2020}
  and randomly sampled designs \cite{nickl2023}, their results do not
  hold for D-optimal designs. Luckily, we can expect convergence from
  D-optimal designs (including clustered designs) in the following
  sense: the posterior uncertainty ellipsoid will contract to zero
  along each one of its eigenvectors. See Section
  \ref{subsub:convergence} for a precise statement and a proof based
  on Theorem \ref{thm:char} --- the main theorem of this manuscript.
\end{quote}
\begin{quote}
  In this section we fulfill our promise from Section
  \ref{subsub:implications} and prove that the posterior uncertainty
  ellipsoid in $\hilo$ will contract to zero along every eigenvector
  of $\fwd \prcov \fwd^*$. In our proof we ignore potential problems
  with conducting inference on function spaces which are not unique to
  D-optimal designs (see e.g.~\cite{owhadi2015} for more details).
  \newline
  \newline  
  First, denote $\opt_m$ a D-optimal design utilizing $m$ measurements
  and denote $k_m := \rank\opt_m^*\opt_m$. An immediate consequence of
  Theorem \ref{thm:char} is that allowing more measurements will
  eventually allow us to measure each eigenvalue, i.e.:
  \begin{equation}\label{eq:lim}
    \lim_{m\to\infty} k_m = \infty
  \end{equation}
  \newline
  \newline
  Now, recall that part (5) of Theorem \ref{thm:char} ensures that the
eigenvalues of the posterior pushforward covariance equal
\begin{equation*}
  \theta^{(k)}_i = \left ( \frac{\sum_{j=1}^{k} \lambda_j^{-1} +
    \sigma^{-2}m}{k} \right )^{-1}, \text{ for $i\leq k$}.
\end{equation*}
Using the inequality for the arithmetic and harmonic means, it is easy
to verify that
\begin{equation*}
  \theta^{(k)}_i \leq \lambda_{k}.
   %% \leq \frac{\sum_{j=1}^{k} \lambda_j}{k}
\end{equation*}
Since $\lim_{k\to \infty} \lambda_k= 0$, we conclude that for all $i$,
$\lim_{k\to\infty} \theta^{(k)}_i = 0$. Combining the latter
observation with eq.~\eqref{eq:lim}, we conclude that
$\lim_{m\to\infty} \theta^{(k_m)}_i = 0$ for all $i$. Therefore,
posterior uncertainty decays to zero for all eigenvectors.
\end{quote}

@@

\AR These are two great points. Indeed, the growth of eigenvalues of
the Laplacian is different in higher dimensions. This is why in higher
dimensions we would have to take as prior \(u_0 \sim
\mathcal{N}(\param_0, (-\Delta)^{-\gamma})\), for some \(\gamma >
d/2\) in order to maintain regularity of prior realizations. I decided
not to include a discussion on the choice of priors in higher spatial
dimensions because (1) it would require too much exposition and make
me digress from the main points I wanted to convey and (2) due to
technical difficulties in installing packages for inverse problems
\cite{attia2023pyoed, villa2021} I had to implement the inverse
problem of the 1D heat equation from scratch, so I decided to leave
out experiments in higher dimensions.

@@

\AR I removed the statement of Theorem 1 from the Introduction. It is
Theorem \ref{thm:char} now.

@@

\begin{quote} %% Replication
  Clusterization should not be confused with replication. Replication
requires that the experimentalist executes multiple trials under
circumstances that are \emph{nominally identical} \cite[Section
  1.2.4]{morris2011}. Replication is commonly viewed as a beneficial
and even necessary aspect of optimal experimental design
\cite{fisher1949design, morris2011, schafer2001replication}.
For example, \cite{fisher1949design}, suggested repeating his famous
milk and tea experiment in order "to be able to demonstrate the
predominance of correct classifications in spite of occasional
errors". Unfortunately, in the experiments we consider, replication is
impossible. For example, in the MRI problem, we cannot generate an
individual nominally identical to the one we wish to scan.
\newline
\newline
Similarly to a design implementing replication, a clustered design
reduces the signal-to-noise ratio of the repeated measurements
\cite{telford2007brief}. The difference is that a clustered design
takes repeated measurements at the expense of other quantities not
measured at all. For example, consider an experiment measuring the
effect of rainfall on grass growth \cite{fay2000rainfall}. The
experiment involved four rainfall manipulation "treatments"
(i.e.~simulating different timing and quantity of rainfall), each
replicated three times over different plots of land. Indeed, it seems
reasonable for researchers to replicate the phenomenon they are trying
to study. A clustered design in such an experiment would imply the
researchers should take a repeated measurement \emph{on the same
plot}, at the expense of measuring grass growth in other plots!
\newline
\newline
To conclude our short discussion on replication vs.~clusterization:
these are fundamentally different concepts and even though replication
is quite intuitive, it is inapplicable to the inverse problems we
consider in this manuscript.
\end{quote}

@@

\AR the Mat\'ern-like prior actually arises by using a Laplacian-like
operator as a prior \cite{rue2011}. So a study of this example (with
and without correlated errors) is already included in the paper. I am
afraid implementing another inverse problem is out of scope for the
present paper since I was not able to install common packages for
inverse problems \cite{attia2023pyoed, villa2021} on my machine.

@@

\AR Indeed, point evaluations $\delta_x$ are not in any function space
I consider in the paper since my analysis is limited to Hilbert
spaces. Point evaluations can be approximated well in
e.g.~$L^2(\Omega)$ so in my opinion this is not a deal breaker. For
example, we can replace point evaluations by $R(x;t) =
t\mathbf{1}_{[-1/2t, 1/2t]}(x)$, where $\mathbf{1}_{[a,b]}$ is an
indicator function on the interval $[a,b]$. We will then need to
change the norm constraints on measurement vectors from unit to
$\|R(\cdot;t)\|^2 = \int R^2(x;t)dx = t$ norm but that does not change
the analysis as long as $t$ is fixed. I barely mentioned this issue in
the paper since I think it is confusing and adds little context. See a
footnote I added:

@@

\AR I added more details to the explaination explain in the revised
manuscript, see Section \ref{subsub:clusterization1}. See text
reproduced below.
\begin{quote}
Consider $\fwd$ and $\prcov$ from \emph{the inverse problem of the
heat equation}. As before, we denote the eigenvalues of
$\fwd\prcov\fwd^*$ by $\lambda_j$. We input these eigenvalues into our
\emph{generic} model, and find a D-optimal design $\opt$ for our
generic model using Theorem \ref{thm:char}. In our generic model, the
measurements we take are best utilized in reducing uncertainty for the
first $k$ eigenvectors. So, a D-optimal design arising from our
\emph{generic model} completely avoids measuring eigenvectors $k+1$
and above.
\newline \newline
Of course, in a real life problem --- such as the inverse problem of
the 1D heat equation --- it is likely impossible to find measurements
for which all eigenvectors $k+1$ and above are zero. However, if the
eigenvalues of $\fwd\prcov\fwd^*$ decay quickly (recall the
square-exponential decay for eigenvalues of the 1D heat equation in
eq.\eqref{eq:decay}), a D-optimal design will try to balance measuring
a small number (i.e.~$k$) of the leading eigenvectors.
%%\end{quote}
%%\begin{quote}
\newline \newline
The abovementioned balance is explored in
Fig.~\ref{fig:eigenvectors}. We allow $m=4$ measurements in $\Omega =
[0,1]$ and observe that D-optimal measurement locations are clustered
at $x_1 = 0.31$ and $x_2 = 0.69$. Upon close inspection of the scaled
eigenvectors of $\fwd \prcov \fwd^*$, we first observe that
eigenvectors $3$ and above have negligible prior amplitude. Since we
only have $m=4$ measurements at our disposal, we interpret these
results, following Theorem \ref{thm:char}, as implying we should only
care about measuring the first and second eigenvectors. Then, we note
the D-optimal $x_1,x_2$ present a compromise between the amplitude of
the first and second eigenvectors. For example, a measurement at
$x=0.5$ would have ignored the second eigenvector altogether, since
the second eigenvector is zero at $x=0.5$.
\newline \newline
Now we can understand measurement clusterization for the inverse
problem of the 1D heat equation. A D-optimal design attempts to
measure the first $k$ eigenvectors of $\fwd \prcov \fwd^*$. But there
may be (spatial) limitations on where these $k$ eigenvectors have
large amplitude. For the inverse problem of the heat equation there
are two spatial locations that present a good compromise between the
amplitudes of the first and second eigenvectors, namely $x_1$ and
$x_2$ --- see Fig.~\ref{fig:eigenvectors}. We have $m=4$ measurements
at our disposal but only two spatial locations that are a good
compromise between the amplitudes of the first and second scaled
eigenvectors. Thus, clusterization arises as a consequence of the
pigeonhole principle.
\end{quote}
\begin{figure}\label{fig:eigenvectors}
    \centering
    \includegraphics[width=\textwidth]{../latex/figs/eigenvectors_dst_scaled.png}
    \caption{D-optimal measurement locations ($m=4$ measurements) and
      weighted eigenvectors for finding the initial condition of the
      1D heat equation. Measurement locations and weighted
      eigenvectors are plotted over the computational domain $\Omega =
      [0, 1]$ (x-axis). Measurement clusterization occurs
      approximately at $0.31$ and $0.69$. These two locations are a
      compromise between the amplitudes of the first and second
      eigenvectors, which are the eigenvectors that a D-optimal design
      aims to measure. Allocating $m=4$ measurements into two
      locations results in clusterization, according to the pigeonhole
      principle.}
  \label{fig:why}
\end{figure}

%% \begin{quote}
%%   Building on Theorem \ref{thm:char}, we can now give a compelling
%%   explanation to the measurement clusterization we observed for the
%%   inverse problem of the heat equation, see Section
%%   \ref{subsub:clusterization1} below. We also suggest a generic
%%   explanation for clusterization, see Section
%%   \ref{subsub:clusterization2}.
%%   \newline
%%   \newline
%%   Consider $\fwd$ and $\prcov$ from \emph{the inverse problem of the
%%   heat equation}. As before, we denote the eigenvalues of
%%   $\fwd\prcov\fwd^*$ by $\lambda_j$. We input these eigenvalues into
%%   our \emph{generic} model, and find a D-optimal design $\opt$ for our
%%   generic model using Theorem \ref{thm:char}. In our generic model,
%%   the measurements we take are best utilized in reducing uncertainty
%%   for the first $k$ eigenvectors. So, a D-optimal design arising from
%%   our \emph{generic model} completely avoids measuring eigenvectors
%%   $k+1$ and above.
%%   \newline
%%   \newline
%%   Of course, in a real life problem --- such as the inverse problem of
%%   the 1D heat equation --- it is likely impossible to find measurement
%%   locations for which all eigenvectors $k+1$ and above are
%%   zero. However, if the eigenvalues of $\fwd\prcov\fwd^*$ decay
%%   quickly (recall the square-exponential decay for eigenvalues of the
%%   1D heat equation in eq.\eqref{eq:decay}), a D-optimal design will
%%   try to balance measuring a small number (i.e.~$k$) of the leading
%%   eigenvectors.
%% \end{quote}
%% \begin{quote}
%%   The abovementioned balance is explored in
%%   Fig.~\ref{fig:eigenvectors}. We allow $m=4$ measurements in $\Omega
%%   = [0,1]$ and observe that D-optimal measurement locations are
%%   clustered at $x_1 = 0.31$ and $x_2 = 0.69$. Upon close inspection of
%%   the scaled eigenvectors of $\fwd \prcov \fwd^*$, we first observe
%%   that eigenvectors $3$ and above have negligible prior
%%   amplitude. Since we only have $m=4$ measurements at our disposal, we
%%   interpret these results, following Theorem \ref{thm:char}, as
%%   implying we should only care about measuring the first and second
%%   eigenvectors. Then, we note the D-optimal $x_1,x_2$ present a
%%   compromise between the amplitude of the first and second
%%   eigenvectors. For example, a measurement at $x=0.5$ would have
%%   ignored the second eigenvector altogether, since the second
%%   eigenvector is zero at $x=0.5$.
%%   \newline
%%   \newline
%%   Now we can understand measurement clusterization for the inverse
%%   problem of the heat equation. A D-optimal design attempts to measure
%%   the first $k$ eigenvectors of $\fwd \prcov \fwd^*$. But there may be
%%   (spatial) limitations on where these $k$ eigenvectors have large
%%   amplitude. For the inverse problem of the heat equation there are
%%   two spatial locations that present a good compromise between the
%%   amplitudes of the first and second eigenvectors, namely $x_1$ and
%%   $x_2$ --- see Fig.~\ref{fig:eigenvectors}. We have $m=4$
%%   measurements at our disposal but only two spatial locations that are
%%   a good compromise between the first and second scaled
%%   eigenvectors. Thus, clusterization arises as a consequence of the
%%   pigeonhole principle.
%% \end{quote}

%% \begin{figure}\label{fig:eigenvectors}
%%   \centering
%%   \includegraphics[width=\textwidth]{../latex/figs/eigenvectors_dst_scaled.png}
%%   \caption{D-optimal measurement locations ($m=4$ measurements) and
%%     weighted eigenvectors for finding the initial condition of the 1D
%%     heat equation. Measurement locations and weighted eigenvectors are
%%     plotted over the computational domain $\Omega = [0, 1]$
%%     (x-axis). Measurement clusterization occurs approximately at
%%     $0.31$ and $0.69$. These two locations are a compromise between
%%     the magnitudes of the first and second eigenvectors, which are the
%%     eigenvectors that a D-optimal design aims to measure. Allocating
%%     $m=4$ measurements into two locations results in clusterization,
%%     according to the pigeonhole principle.}
%%   \label{fig:why}
%% \end{figure}
